\documentclass{article}
\usepackage{graphicx} % Required for inserting images

\title{Power-Optimal Inverter Chain Design Under Elmore Delay Constraints}
\author{Suprith Krishnakumar, John Fang, John Hsu, Yuheng Li}
\date{March 2026}

\begin{document}

\maketitle

\section{Introduction}
\subsection{Motivation}
	Modern VLSI digital circuits demand increasingly stringent performance 
	and energy efficiency requirements, particularly as transistor scaling continues 
	to push designs toward higher density and faster clock speeds. A fundamental 
	component in high-speed digital design is the use of inverter chains (buffer chains) 
	to drive large capacitive loads, where improper sizing or staging can lead to excessive 
	delay or unnecessary power consumption. While classical design heuristics such as fixed 
	stage effort or logical effort methods provide useful guidelines, they do not always yield 
	globally optimal solutions under complex constraints involving wire parasitics, transistor 
	sizing limits, and timing requirements. The goal of this work is to develop a principled 
	optimization framework for inverter chain design that minimizes dynamic power consumption 
	subject to an Elmore delay constraint. We approach this problem through the lens of convex 
	optimization by reformulating the inherently non-convex delay expressions into a geometric 
	programming (GP) framework via logarithmic transformation. This enables efficient computation 
	of globally optimal inverter sizes and interconnect lengths while accounting for physical 
	constraints. Through this formulation, we aim to bridge classical circuit design intuition 
	with modern optimization techniques and provide a scalable method for a energy-efficient 
	buffer chain design.
\subsection{Previous Works}

\subsection{Intended Contributions}
\begin{enumerate}
	\item Optimization-Based Formulation: We formulate the inverter chain design problem as a 
	power minimization problem with Elmore delay, sizing, and interconnect constraints.
	\item Geometric Programming Reformulation: We show that the non-convex delay constraint can 
	be expressed as a posynomial and transformed into a convex constraint via logarithmic variable 
	substitution.
	\item Convexity Verification: We provide a formal justification that the transformed delay 
	constraint is convex by leveraging the log-sum-exp structure and verifying all terms satisfy 
	geometric programming requirements.
	\item Solver Implementation: We implement the resulting formulation in a convex optimization 
	framework (e.g., CVXPY) to compute optimal inverter sizing and interconnect parameters.
	\item Design Insight: We analyze the trade-offs between delay, sizing, and power, and compare 
	the optimized solutions to classical inverter chain design heuristics.
\end{enumerate}
\subsection{Organization of the Paper}


\section{Statement of the Problem}
\subsection{Primal Formulation}
\subsubsection{Physical Interpretation of the Inverter Chain}
We consider a chain of $N$ CMOS inverters used to drive a fixed load capacitance $C_L$. 
Such inverter chains, commonly referred to as buffer chains, are fundamental building 
blocks in high-speed digital circuits, where they are used to drive large capacitive 
loads while meeting strict timing constraints.

Each stage in the inverter chain is characterized by two primary design variables:
\begin{itemize}
    \item $W_i > 0$: the effective transistor width (or sizing factor) of inverter $i$
    \item $l_i > 0$: the interconnect length between inverter $i$ and inverter $i+1$
\end{itemize}
The objective is to determine the optimal values of $\{W_i\}_{i=1}^N$ and $\{l_i\}_{i=1}^N$ 
that minimize dynamic power consumption while satisfying a delay constraint.
Each inverter contributes capacitance from two sources:
\paragraph{Gate (Input) Capacitance}
The input capacitance of inverter $i$ is proportional to its size:
\[ C_{in,i} = W_i C_{g0} \]
where $C_{g0}$ represents the gate capacitance of a unit-sized inverter.

\paragraph{Parasitic Capacitance}
Each inverter also has intrinsic parasitic capacitance due to diffusion regions:
\[ C_{p,i} = W_i C_{p0} \]
where $C_{p0}$ is the parasitic capacitance of a unit inverter.

These capacitances contribute to both dynamic power consumption and delay.
The wires connecting inverter stages introduce additional parasitics:
\paragraph{Interconnect Modeling}
\begin{itemize}
    \item Resistance: \[ R_{wire,i} = r\,l_i \]
    \item Capacitance: \[ C_{wire,i} = c\,l_i \]
\end{itemize}
where $r$ and $c$ denote per-unit-length resistance and capacitance, respectively. 
These parameters capture the physical effects of interconnect scaling and play a 
critical role in delay modeling.

\paragraph{Inverter Resistance}
The effective output resistance of inverter $i$ is inversely proportional 
to its size: \[ R_i = \frac{R_0}{W_i} \] where $R_0$ is the resistance of a unit 
inverter. Increasing the transistor width reduces resistance, improving delay at the 
expense of increased capacitance.

\paragraph{Dynamic Power Objective}
The dynamic power consumption of a digital circuit is given by:
\[ P_{dyn} = \alpha f V_{DD}^2 C_{tot} \]
where $\alpha$ is the switching activity factor, $f$ is the clock frequency, and $V_{DD}$ is the supply voltage. 
These quantities are treated as constants in our formulation.

Thus, minimizing dynamic power reduces to minimizing the total switched capacitance $C_{tot}$.

The total capacitance in the inverter chain consists of:
\begin{itemize}
    \item Parasitic capacitance of each inverter
    \item Gate capacitance of downstream inverters
    \item Interconnect capacitance
    \item Final load capacitance $C_L$
\end{itemize}

This yields:
\[ C_{tot} = \sum_{i=1}^{N} W_i C_{p0} + \sum_{i=1}^{N-1} W_{i+1} C_{g0} + \sum_{i=1}^{N} c\,l_i + C_L \]
Since $C_L$ is constant, it does not affect the optimization and may be omitted from the objective.

\paragraph{Elmore Delay Constraint}

To ensure correct circuit operation, the total propagation delay must not exceed a specified clock period $T_{clk}$. 
We model delay using the Elmore delay approximation, which expresses delay as a weighted sum of downstream capacitances.
The delay contribution of each stage includes interactions between inverter resistance and downstream capacitances, 
as well as interconnect effects.

For stages $i = 1, \dots, N-1$, the delay contribution is:
\[ d_i = \; R_0 C_{g0}\frac{W_{i+1}}{W_i} + R_0 C_{p0} + R_0 c\frac{l_i}{W_i} \\
+ r C_{g0} l_i W_{i+1} + r C_{p0} l_i W_i + r c l_i^2 \]

The final stage differs since it drives the external load $C_L$:
\[ d_N = \; R_0 C_L\frac{1}{W_N} + R_0 C_{p0} + R_0 c\frac{l_N}{W_N} \\
+ r C_L l_N + r C_{p0} l_N W_N + r c l_N^2 \]
The total delay constraint is therefore: \[ \sum_{i=1}^{N-1} d_i + d_N \le T_{clk} \]

\paragraph{Design Constraints}
To ensure physically realizable designs, we impose:
\[ W_{min} \le W_i \le W_{max}, \quad W_i > 0, \quad l_i > 0 \]

\subsubsection{Primal Statement}
\paragraph{Geometric Programming Reformulation}
The original formulation is non-convex due to multiplicative and ratio terms such as:
\[ \frac{W_{i+1}}{W_i}, \quad W_i l_i, \quad l_i^2 \]
To enable efficient optimization, we reformulate the problem using geometric programming techniques.

\paragraph{Logarithmic Variable Transformation}
We introduce the change of variables:
\[ x_i = \log W_i, \qquad y_i = \log l_i \]
which implies:
\[ W_i = e^{x_i}, \qquad l_i = e^{y_i} \]
This transformation is bijective over the positive orthant, ensuring no loss of generality.

\paragraph{Transformation of Nonlinear Terms}
Under this transformation, multiplicative and ratio terms become exponentials of affine expressions:
\[
\frac{W_{i+1}}{W_i} = e^{x_{i+1}-x_i}, \quad
W_i l_i = e^{x_i+y_i}, \quad
l_i^2 = e^{2y_i}
\]
Thus, each term in the delay expressions can be written in the form:
\[ c_k e^{a_k^T z}, \quad \text{where } z = [x_1, \dots, x_N, y_1, \dots, y_N]^T \]

\paragraph{Convexity of the Log-Sum-Exp Function}
Define:
\[ f(x) = \log \left( \sum_{k=1}^{n} e^{x_k} \right) \]
Let $z_k = e^{x_k}$ and $z = [z_1, \dots, z_n]^T$. Then:
\[ \nabla^2 f(x) = \frac{1}{\mathbf{1}^T z} \mathrm{diag}(z) - \frac{1}{(\mathbf{1}^T z)^2} zz^T \]
For any $v \in \mathbb{R}^n$:
\[
v^T \nabla^2 f(x) v
= \frac{1}{(\mathbf{1}^T z)^2}
\left[
(\mathbf{1}^T z)\left(\sum_{i=1}^n z_i v_i^2\right)
- \left(\sum_{i=1}^n z_i v_i\right)^2
\right]
\]
By the Cauchy–Schwarz inequality:
\[ \left(\sum_{i=1}^n z_i v_i\right)^2 \le (\mathbf{1}^T z)\left(\sum_{i=1}^n z_i v_i^2\right) \]
Thus:
\[ v^T \nabla^2 f(x) v \ge 0 \quad \forall v \]
which proves that $f(x)$ is convex.

\paragraph{Convexity of the Delay Constraint}
The total delay can be written as:
\[ D(x,y) = \sum_k c_k e^{a_k^T z} \]
Taking logarithms:
\[ \log D(x,y) = \log \left( \sum_k e^{a_k^T z + \log c_k} \right) \]
which is a log-sum-exp function and therefore convex. Hence, the constraint:
\[ D(x,y) \le T_{clk} \]
is equivalent to the convex constraint:
\[ \log D(x,y) \le \log T_{clk} \]

\paragraph{Final Convex Optimization Problem}
The optimization problem becomes:
\[ \min_{x,y} \quad & \log \left(
C_{p0} e^{x_1} + \sum_{i=2}^{N} (C_{g0}+C_{p0}) e^{x_i}
+ c \sum_{i=1}^{N} e^{y_i} \right) \]
\[ \text{s.t.} \quad &
\log \left( \sum_{i=1}^{N-1} d_i(x,y) + d_N(x,y) \right)
\le \log T_{clk} \] 
\[ \log W_{min} \le x_i \le \log W_{max}, \quad \forall i \]
\[ \log \epsilon \le y_i \le \log l_{max}, \quad \forall i \]

\paragraph{Strong Duality and Slater's Condition}
The problem is convex since:
\begin{itemize}
    \item The objective is a log-sum-exp function, hence convex.
    \item The delay constraint is convex.
    \item The bound constraints are affine.
\end{itemize}

Let the feasible set be:
\[
\mathcal{D} =
\left\{
(x,y) \mid
\log W_{min} \le x_i \le \log W_{max},\;
\log \epsilon \le y_i \le \log l_{max}
\right\}
\]

The relative interior of $\mathcal{D}$ is:
\[
\mathrm{relint}(\mathcal{D}) =
\left\{
(x,y) \mid
\log W_{min} < x_i < \log W_{max},\;
\log \epsilon < y_i < \log l_{max}
\right\}
\]

To verify Slater’s condition, we must show that there exists
$(x^\ast, y^\ast) \in \mathrm{relint}(\mathcal{D})$ such that:
\[
\log \left( \sum_{i=1}^{N-1} d_i(x^\ast,y^\ast) + d_N(x^\ast,y^\ast) \right)
< \log T_{clk}
\]

Such a point exists by choosing:
\begin{itemize}
    \item $y_i^\ast$ sufficiently small (close to $\log \epsilon$), which reduces all terms involving $e^{y_i}$,
    \item $x_i^\ast$ strictly within bounds and arranged to balance exponential terms.
\end{itemize}

Since all delay terms are positive and continuous, and can be made sufficiently small relative to $T_{clk}$, a strictly feasible point exists.

Therefore, Slater’s condition holds:
\[ \exists (x^\ast, y^\ast) \in \mathrm{relint}(\mathcal{D}) \] 
such that \[  f_i(x^\ast,y^\ast) < 0 \]

\paragraph{Conclusion}
Since the problem is convex and Slater’s condition holds, strong duality follows:
\[
p^\ast - d^\ast = 0
\]
Thus, the duality gap is zero, and the Karush-Kuhn-Tucker (KKT) conditions are necessary and sufficient for optimality.


\subsection{Dual Formulation}

\subsection{KKT Conditions}


\section{Approaches}

\section{Results And Conclusion}
\subsection{Results}

\subsection{Conclusion}


\section{References}


\end{document}
